---
title: "Titanic Death Rates"
subtitle:
author: "Cade, Zety, Kristen and Andrew"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# C:/Users/andre/Desktop/presentation/presentation_files/Datasets/titanic.csv
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "img/Titanic.jpg"
)
```

class: center, middle

## Our Research Question:
## How does the different variables of the passengers on the Titanic affect their survival rates?

---

class: inverse, center, middle

# Data Information and Our Thoughts on What We Think the Data Will Show

---

# Where Did we Aquire the Data 
- Firstly we searched through the links provided on the IDS workshop page

- We then investigated the different link looking through the different data sets 

- We then found our data set by going through https://github.com/awesomedata/awesome-public-datasets

- From this we selected Titanic Survival Data Set (https://www.kaggle.com/c/titanic/data)

- From this we downloaded the Data into R so we begin working on it.

---

class: inverse, middle, center

# What are the different variables?

---

# The different variables

- survival (1=Yes; 0=No)
- pclass = Ticket Class (1=first class; 2=second class; 3=third class)
- sex
- age = Age in years
- sibsp = number of siblings / spouses aboard the Titanic
- parch = number of parents / children aboard the Titanic
- ticket = Ticket number
- fare = Passenger fare
- cabin = Cabin number
- embarked = Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton)

---

# Variables Notes

- age:
fractional if <1
if estimated in the form of xx.5
- sibsp:
sibling = brother/sister/stepbrother/stepsister
spouse = husband/wife (mistresses/fiancés ignored)
- parch: 
parent = mother/father
child = daughter/son/stepdaughter/stepson
for nannies, parch=0

---

# Background on the Titanic and Our Thoughts

-On April 15, 1912, during her maiden voyage, the widely considered “unsinkable” RMS Titanic sank after colliding with an iceberg. Unfortunately, there weren’t enough lifeboats for everyone on-board, resulting in the death of 1502 out of 2224 passengers and crew. Resulting in a 67.5% survival rate out of the passengers 

-While there was some element of luck involved in surviving, it seems some groups of people were more likely to survive than others.

-We believe that people who were female, with a younger age and higher class will have a higher survival rate of surviving the Titanic ship wreck.

-We think this as the younger age means they should be able to deal with the harsh conditions better and them being better class means they would be located in better structurally built part of the ship.

-Being a female, we think they'd get priority when boarding the lifeboats therefore having a higher chance of survival  

---

```{r}
titanic <- read_csv("/Users/kristen/Documents/HW/Project/Datasets/titanic.csv") 
```

---

```{r}
glimpse(titanic)
```

---

# Any statistics that you believe will be useful in answering your question(s)?

BAR PLOT - number of people survived against factor (sex, class, age)
- each factor have its own subset
- allow us to compare the data (does class plays a role in the survival of the passengers? or age or sex)
- if they do, then which factor contributes the most (face wrap? // stacked bar chart // multiple axis chart)

Pie chart, line plot, box plot

---

# Exploratory Data Analysis

What is Exploratory Data Analysis?
Exploratory Data Analysis, also known as Data Exploration, is a step in the Data Analysis Process, where a number of techniques are used to better understand the dataset being used.

‘Understanding the dataset’ can refer to a number of things including but not limited to:

1) Extracting important variables and leaving behind useless variables

2) Identifying outliers, missing values, or human error

3) Understanding the relationships, or lack of, between variables

Ultimately, maximizing your insights of a dataset and minimizing potential error that may occur later in the process.

---

class: middle

Within the dataset, titanic, some variables are useless such as "Cabin" and "Ticket" as these variables produce no important information that can be analaysed. They can not be summarised or made into a useful visualisation. 

Variables such as "Age", "Pclass", "Sex", "Survived" and Fare are the most useful vairables as they can be used in both an analytical and predictive way. These variables can be used to identify trends while they can also provide more relevant insights by visualising, transforming and modelling the data. In the dataset, there are numerous missing values such as "Age". This could be due to the failure in observations or that the ages simply where never known.

Through analysing the dataset as a whole and its variables individually, we will gain a better insight of the data through understanding relationships between variables. From quick glance, it is already evident that the passengers who survived where based on their sex, age and class. This will be further explored throughout the presentation.

---

```{r}
my_data <- as_tibble(titanic)
my_data
```

---

```{r}
my_data %>%
  summarise(mean_fare = mean(Fare, na.rm = TRUE))
```

```{r}
my_data %>% 
  summarise(mean_Pclass = mean(Pclass, na.rm = TRUE))
```

---

```{r}
my_data %>% 
  summarise(mean_age = mean(Age, na.rm = TRUE))
```

Through using summary statistics, the dataset is summarised and information is provided. It outlines something about the values in the dataset such as the mean of specific variables (as shown above).

It simplifies the data in order to quickly identify the information needed while also allowing us to identify what looks 'normal' and what doesn't. This outlines potential anomalies. The summary statistics are filled with information and insights extracted from the table. 

---

```{r}
titanic %>%
group_by(Pclass, Survived) %>%
  summarise(
    n = n())
```

Within the summary statistic above, we can see that the variables "Pclass" and "Survived" were grouped in order to find how many people survived based on their class. The dataset is broken down in a way that is easy to read but it is also an efficient way to gather the required information. Through using the function 'summarise(n = n())' the values for who survived is automatically calculated which therefore saves having to count manually.

---

```{r}
na.omit(titanic) %>%
  summarise(titanic,
            min_Age = min(Age),
            max_Age = max(Age),
            min_Fare = min(Fare),
            max_Fare = max(Fare))
```

---

```{r warning=FALSE, out.width="100%", fig.width=6, echo=FALSE}
ggplot(titanic, aes(y = Pclass,
                    fill = Survived)) +
  geom_bar(fill = "lightblue", colour = "Blue") +
  labs(x = "Number of Passengers who Survived",
       y = "Class",
       title = "The Number of Passengers who Survived Based on Class")
```

---

class: middle, center

A bar chart is used when you want to show a distribution of data points or perform a comparison of metric values across different subgroups of your data. From a bar chart, we can see which groups are highest or most common, and how other groups compare against the others. Since this is a fairly common task, bar charts are a fairly ubiquitous chart type.

The data represents the number of passengers who survived based on their class. Through the appropriate use and labelling of the x and y axis, it is clear that most passengers who survived were in third class.

---

```{r warning=FALSE, out.width="100%", fig.width=6, echo=FALSE}
ggplot(data = titanic,
       mapping = aes(x = Survived, y = Age)) +
  geom_boxplot () +
  facet_wrap(~Pclass) +
   labs(x = "Number of Passengers who Survived",
        title = "The Number of Passengers who Survived Based on Class and Age")
```

---

class: middle, center

The data represents the number of passengers who survived based on class and age. The variable "Pclass" is split into a readable graph by the function facet wrap, splitting the data on the vertical grid to distinguish between each class. There is a direct relation between class, age and who survived which forms the structure of the graph e.g. the age of those who survived in first class. The graph is in the form of a boxplot which showcases the maximum and minimum age of the survivors as well as the median and the outliers showcasing unique ages. The boxplot is easy to read and follow therefore valuing the respects of visualisation.

---

```{r warning=FALSE, out.width="100%", fig.width=6, echo=FALSE}
ggplot(titanic, aes(x = Age)) +
  geom_density() +
  labs(x = "Age", title = "Passengers who Survived based on Age and Sex") +
  facet_wrap(~Sex, nrow = 2)
```

---

class: middle, center

The data represents the number of passengers who survived based on age and sex. Facet wrap splits the data on the into two rows to distinguish between each sex. There is a direct relation between age, sex and who survived which forms the structure of the graph e.g. the sex and age of those who survived. The graph is in the form of a densityplot which showcases clearly each variable and its trends. 

---

```{r warning=FALSE, out.width="100%", fig.width=6, echo=FALSE}
ggplot(titanic, aes(y = Pclass)) +
  facet_grid(~Sex) +
  geom_bar(fill = "lightblue", color = "blue") +
  labs(x = "Number of Passengers who Survived",
       y = "Class",
       title = "The Number of Passengers who Survived Based on Class and Sex")
```

---

class: middle, center

The data represents the number of passengers who survived based on class and sex. Facet wrap splits the data on the vertical grid to distinguish between each sex. There is a direct relation between class, sex and who survived which forms the structure of the graph e.g. the sex of those who survived in third class. The graph is in the form of a barplot which outlines clearly each variable and its values in a clear, user-friendly way. Like the boxplot, the barplot is easy to read and follow therefore valuing the respects of visualisation.

---

# What results from these statistics are needed to support your hypothesized answer?

- no hypothesized answer currently
- but if there is eg: class does play a role in the passengers' survival rate, then the data visualization should show a significant difference in the number of ppl survived per class
- another example: women and children were prioritized during the titanic rescue mission, then we should see in the data that the number of women and children will be significantly more than the rest

---
class: inverse, middle, center

# Any Questions?

---

# Code

```{r boring-analysis}
# a boring analysis
ave <- mean(cars$dist)
print(ave)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

ggplot(titanic, mapping =  aes(x = age, 
  fill = sex) +
  geom_bar()
  labs(
  x= age
  y= number 
  title = Age and Sex of Passengers on the Titanic
)

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
